---
title: "Egonetwork example"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.align='center')
```

# Load the data

```{r packages,warning=FALSE,message=FALSE}
library(igraph)
library(graphlayouts)
library(ggraph)
#remotes::install_github("thomasp85/patchwork)
library(patchwork)
```

```{r load_data}
ego_files <- list.files("../data/egonet/",full.names = T)
egonets <- lapply(ego_files,function(x) read.graph(x,format="graphml"))
```

`egonets` is a list of 32 ego networks. You can access them individually as follows.

```{r egonet_ex}
egonets[[1]]
```

Each network has three node variables(`age`, `rank` and `gender`) and one edge variable (`weight`).
You can get a quick view of the network via `qgraph`.

```{r quickviz,fig.width=4,fig.height=4}
qgraph(egonets[[1]])
```

# radial layout

For ego networks, a radial layout is appropriate, for instance, when there exists a 
node attribute that ranks the alters proximity to ego. This attribute can be used
with `layout_with_centrality()` to create an ego-centric layout. Note, however, that 
centraity based layouts put nodes with higher values in the center. We thus need to invert the
ranks given in the data (rank 1 means very close to ego).

The figure below shows the basic radial layout plot without any additional informations.
```{r radial_ego1}
ggraph(egonets[[1]],layout = "centrality", cent = 5-V(egonets[[1]])$rank)+
  geom_edge_link0()+
  geom_node_point()+
  theme_graph()
```

In the next step, we add some more aesthetics to the plot. We map the node fill color to 
gender and edge width to the weight attribute

```{r radial_ego2}
ggraph(egonets[[1]],layout = "centrality", cent = 5-V(egonets[[1]])$rank)+
  geom_edge_link0(aes(edge_width = weight))+
  geom_node_point(aes(fill=gender),shape = 21,size = 5)+
  theme_graph(base_family = "Helvetica")
```

This is already quite ok, though edge are clearly way to thick. We also change the 
default colors of the nodes and edges in the code below. 

```{r radial_ego3}
ggraph(egonets[[1]],layout = "centrality", cent = 5-V(egonets[[1]])$rank)+
  geom_edge_link0(aes(edge_width = weight),edge_colour="grey66")+
  geom_node_point(aes(fill = gender),shape = 21,size = 5)+
  scale_edge_width_continuous(range = c(0.2,1.2))+
  scale_fill_manual(values = c("w" = "#EEB422", "m" = "#3D3D3D"))+
  theme_graph(base_family = "Helvetica")

```

As a last step, we can add optional circles to the plot to emphasize the position of nodes.
Note that we draw them **before** edges and nodes, so that they are in the background.

```{r radial_ego4}
ggraph(egonets[[1]],layout = "centrality", cent = 5-V(egonets[[1]])$rank)+
  draw_circle(use = "cent")+
  geom_edge_link0(aes(edge_width = weight),edge_colour="grey66")+
  geom_node_point(aes(fill = gender),shape = 21,size = 5)+
  scale_edge_width_continuous(range = c(0.2,1.2))+
  scale_fill_manual(values = c("w" = "#EEB422", "m" = "#3D3D3D"))+
  theme_graph(base_family = "Helvetica")
```

The circles look a little too elliptic. We can fix the aspect ratio with `coord_fixed`.
Additionally, we do not need the legend for edge widths and turn it of with `guide = FALSE`
in `scale_edge_width_continuous`. The legend for the node colors is placed at the bottom.

```{r radial_ego5}
ggraph(egonets[[1]],layout = "centrality", cent = 5-V(egonets[[1]])$rank)+
  draw_circle(use = "cent")+
  geom_edge_link0(aes(edge_width = weight),edge_colour="grey66")+
  geom_node_point(aes(fill = gender),shape = 21,size = 5)+
  scale_edge_width_continuous(range = c(0.2,1.2),guide = FALSE)+
  scale_fill_manual(values = c("w" = "#EEB422", "m" = "#3D3D3D"))+
  coord_fixed()+
  theme_graph(base_family = "Helvetica")+
  theme(legend.position = "bottom")
  
```

There are still some flaws in the plot (overlapping nodes and ego might not be relevant to show), 
but overall it is a nice way of plotting an ego network using the `graphlayouts` package.

# Small multiples

In some cases, it might be needed to show a large collection of ego networks in one plot.
We can easily reuse the code from above for our entire ego network collection that is
stored in `egonets`. We start by turning the ggraph code into a function, which only takes
a network as input.

```{r ego_fct}
plot_ego <- function(net){
  ggraph(net,layout = "centrality", cent = 5-V(net)$rank)+
  draw_circle(use = "cent")+
  geom_edge_link0(aes(edge_width = weight),edge_colour="grey66")+
  geom_node_point(aes(fill = gender),shape = 21,size = 5)+
  scale_edge_width_continuous(range = c(0.2,1.2),guide = FALSE)+
  scale_fill_manual(values = c("w" = "#EEB422", "m" = "#3D3D3D"))+
  coord_fixed()+
  theme_graph(base_family = "Helvetica")+
  theme(legend.position = "bottom")
}
```

Using `lapply`, we apply it to our whole collection.
```{r ego_plots}
ego_plots <- lapply(egonets,plot_ego)
```

You can now look at specific networks with

```{r ego_plots1}
ego_plots[[12]]
```

or we can stitch them all together with the `patchwork` package. Doing so is pretty straightforward.
```{r ego_plots2}
ego_plots[[1]] + ego_plots[[2]]
```

\newpage

To put all networks into one final plot, we use a for loop.

```{r ego_plotsall}
p <- ego_plots[[1]]
for(i in 2:32){
 p <- p + ego_plots[[i]] 
}
```

The patchwork package has some additional functions which allow us to, e.g., control 
how many single plots are put into a row.

```{r ego_plotsshow,fig.width=20,fig.height=14}
p + plot_layout(ncol=8)
```


